Air Quality Data - PM2.5 Components

### 2021-01-11 code/data-processors/air-data-processor.Rmd
added daily PM data based on closest zipcode

```{r load data}
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\state-region.RDa')
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\site-zip.RDa')
```

@@

# find the diff bw spec & improve
monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.imprv <- dt.imprv.raw %>% distinct(Monitor.ID) 
monitor.diff <- setdiff(monitor.imprv$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame() 
colnames(monitor.diff) <- 'Monitor.ID'
# improve has 32 Monitor.ID that spec does not have

site.spec <- dt.spec.w %>% distinct(Site.ID)
site.imprv <- dt.imprv.raw %>% distinct(Site.ID) 
site.diff <- setdiff(site.imprv$Site.ID, site.spec$Site.ID) %>% 
  as.data.frame()
colnames(site.diff) <- 'Site.ID'
# improve has 28 Site.ID that spec does not have

dt.new <- merge(dt.imprv.raw, monitor.diff)
dt.pmc.w <- rbind(dt.spec.w, dt.new) %>% 
  arrange(Site.ID, POC, Date)

@@

monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.csn <- dt.csn.raw %>% distinct(Monitor.ID) 
monitor.diff <- setdiff(monitor.csn$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame()
# no difference between two dataset!
# csn is not adding any new information to spec
```

@@

#improve sites added to speciation sites
dt.site.carbon.method <- site.diff %>% 
  mutate(Method = "EC.TOT-OC.TOR") %>% 
  add_row(dt.site.carbon.method) %>% 
  arrange(as.character(Site.ID))

@@

dt.site.carbon.method %>% group_by(Method) %>% 
  summarise(n = n())
```

```{r Air quality data Process}
map(dt.pmc.w, ~sum(is.na(.)))

#sites with measurements for all components - spec & improve
dt.pmc.c <- dt.pmc.w[complete.cases(dt.pmc.w),] 
monitor.pmc <- dt.pmc.c %>% distinct(Monitor.ID) # 698 -> 530 
site.pmc <- dt.pmc.c %>% distinct(Site.ID) #574 -> 486

@@

dt.pmc.sel <- merge(dt.pmc.sel, site.POC.larger) 
monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 698 -> 530 -> 346
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 486 -> 346
dt.pmc.sel$POC <- dt.pmc.sel$Monitor.ID <- NULL
save(dt.pmc.sel, file = here('data', 'processed', 'pmc-selected.RDa'))
```

```{r Smoothing Data & Caluculating 1-yr mvavg for each month}
dt.air <- dt.pmc.sel %>% 
  pivot_longer(cols = AS:OC,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

@@

dt.site.pmc <- dt.pmc.1yr %>% distinct(Site.ID) %>%  #574 -> 486 -> 346 -> 346
  merge(dt.aqs.sites)
save(dt.site.pmc, file = here('data', 'processed', 'site-pmc.RDa'))

dt.site.pmc %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
#Location.Setting  n
#                  2			
#RURAL	           202			
#SUBURBAN	         69			
#URBAN	           73

dt.pmc.1yr %>% 
  distinct(Site.ID, Year) %>% 
  group_by(Site.ID) %>% 
  summarise(n = n()) %>% filter(n > 3) %>% 
  nrow() #346: all  #339: with 4+ years
```

@@

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz6 %>% distinct(Zip.Code) %>% nrow() #4351
dt.pmc.site.zip.bz6 %>% distinct(Site.ID) %>% nrow() #346 > 277

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz12 %>% distinct(Zip.Code) %>% nrow() #8454
dt.pmc.site.zip.bz12 %>% distinct(Site.ID) %>% nrow() #346 > 328

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz24 %>% distinct(Zip.Code) %>% nrow() #15856
dt.pmc.site.zip.bz24 %>% distinct(Site.ID) %>% nrow() #346 > 341

>>>>

```{r load data}
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\state-region.RDa')
```

@@

# find the diff bw spec & improve
monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.imprv <- dt.imprv.raw %>% distinct(Monitor.ID) 
monitor.diff.imprv <- setdiff(monitor.imprv$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame() 
colnames(monitor.diff.imprv) <- 'Monitor.ID'
# improve has 32 Monitor.ID that spec does not have

site.spec <- dt.spec.w %>% distinct(Site.ID)
site.imprv <- dt.imprv.raw %>% distinct(Site.ID) 
site.diff.imprv <- setdiff(site.imprv$Site.ID, site.spec$Site.ID) %>% 
  as.data.frame()
colnames(site.diff.imprv) <- 'Site.ID'
# improve has 28 Site.ID that spec does not have
```

@@

monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.csn <- dt.csn.raw %>% distinct(Monitor.ID) 
monitor.diff.csn <- setdiff(monitor.csn$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame()
# no difference between two dataset!
# csn is not adding any new information to spec
```

```{r add new component data to epa data}
# Only IMROVE had new data
# CSN was same as epa
dt.new <- merge(dt.imprv.raw, monitor.diff.imprv)
dt.pmc.w <- rbind(dt.spec.w, dt.new) %>% 
  arrange(Site.ID, POC, Date)

@@

#improve sites added to speciation sites
dt.site.carbon.method <- site.diff.imprv %>% 
  mutate(Method = "EC.TOT-OC.TOR") %>% 
  add_row(dt.site.carbon.method) %>% 
  arrange(as.character(Site.ID))

@@

dt.site.carbon.method %>% group_by(Method) %>% 
  summarise(n = n())
# TOR 167
# TOT 326
# EC.TOT-OC.TOR 28
```

```{r Adding Daily PM data to pmc daily data}
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\site-zip-closest-bz12.RDa')

# closest zipcode to each site in each year in 12-mile bufferzone
dt.site.close.zip.pmc <- site.pmc %>% 
  merge(dt.site.zip.closest.bz12) %>% 
  arrange(Site.ID, Year)

save(dt.site.close.zip.pmc, file = here('data', 'processed', 'site-close-zip-pmc.RDa'))
write.csv(dt.site.close.zip.pmc, file = here('data', 'processed', 'site-close-zip-pmc.csv'))

dt.site.close.zip.pmc %>% distinct(Site.ID) %>% nrow() #560
# 560 out of 574 sites have zipcodes in their 12-mile buffer

# site-close-zip-pmc.csv was uploaded on NEU Cluster to be merged with daily zipcode level
# ... PM data.

# daily PM value for each site based on closest zipcode value
dt.pm.daily <- read.csv(here('data','processed','pm_pmc_daily.csv')) %>% 
  transmute(Site.ID = Site_ID,
            Date = mdy(chardate),
            PM)

dt.pmc.pm.w <- dt.pmc.w %>% 
  merge(dt.pm.daily)

save(dt.pmc.pm.w, file = here('data', 'processed', 'pmc-pm-wide.RDa'))
```

```{r Air quality data Process}
map(dt.pmc.pm.w, ~sum(is.na(.)))

#sites with measurements for all components - spec & improve
dt.pmc.c <- dt.pmc.pm.w[complete.cases(dt.pmc.pm.w),] 
monitor.pmc <- dt.pmc.c %>% distinct(Monitor.ID) # 698 -> 530  -> 515
site.pmc <- dt.pmc.c %>% distinct(Site.ID) #574 -> 486 -> 471

@@

dt.pmc.sel <- merge(dt.pmc.sel, site.POC.larger) 
monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 698 -> 530 -> 346 -> 318
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 486 -> 346 -> 318
dt.pmc.sel$POC <- dt.pmc.sel$Monitor.ID <- NULL
save(dt.pmc.sel, file = here('data', 'processed', 'pmc-selected.RDa'))
```

```{r Smoothing Data & Caluculating 1-yr mvavg for each month}
dt.air <- dt.pmc.sel %>% 
  pivot_longer(cols = AS:PM,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

@@

dt.site.pmc <- dt.pmc.1yr %>% distinct(Site.ID) %>%  #574 -> 486 -> 346 -> 318
  merge(dt.aqs.sites)
save(dt.site.pmc, file = here('data', 'processed', 'site-pmc.RDa'))

dt.site.pmc %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
#Location.Setting  n
#                  2			
#RURAL	           177			
#SUBURBAN	         67			
#URBAN	           72

dt.pmc.1yr %>% 
  distinct(Site.ID, Year) %>% 
  group_by(Site.ID) %>% 
  summarise(n = n()) %>% filter(n > 3) %>% 
  nrow() #318: all  #311: with 4+ years
```

@@

load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\site-zip.RDa')

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz6 %>% distinct(Zip.Code) %>% nrow() #4253
dt.pmc.site.zip.bz6 %>% distinct(Site.ID) %>% nrow() #318 > 270

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz12 %>% distinct(Zip.Code) %>% nrow() #8313
dt.pmc.site.zip.bz12 %>% distinct(Site.ID) %>% nrow() #318 > 316

@@

# Count number of zipcodes and sites
dt.pmc.site.zip.bz24 %>% distinct(Zip.Code) %>% nrow() #15601
dt.pmc.site.zip.bz24 %>% distinct(Site.ID) %>% nrow() #318 > 317
